---
title: "Non-Homebased Trips"
author: "Caliper Corporation"
date: "September 30, 2024"
output: 
  html_document:
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(dplyr.summarise.inform = FALSE)
options(scipen = 999)

library(tidyverse)
library(knitr)
library(kableExtra)
library(sf)
source("R/estimate_nhb.R")
source("R/utils.R")
```

## Introduction

There are many problems related to non-home-based trips in traditional
trip-based models arising from the fact that they are disconnected from the
home-based trips with which they comprise complete tours. In order to properly
represent non-home-based trips, two spatial distribution or destination/spatial
choice models are required to account for both the trip’s origin location and
destination location. The four-step model architecture is fundamentally flawed
because it produces non-home-based trips from only one trip distribution or
spatial choice model. 

To address these problems, the model adopts an alternative approach with a simple
change to the structure of the trip-based model, running the non-home-based
model components after and conditional on the home-based model components
instead of in parallel and independently of them as in the traditional four-step
model. This relatively simple structural change significantly improves the model’s
ability to represent non-home-based trips and their response to land use changes
and transportation infrastructure investments. Running a NHB distribution or
destination choice model after and conditional on home-based destination choices
in this approach, provides the required second spatial distribution model to
properly model both the origin and destination of NHB trips.

In this approach, NHB trips are generated separately by mode based on home-based
mode choices. This essentially provides information about whether a traveler has
a car with them and allows the model, despite its trip-based form, to ensure a
reasonable consistency of modes on tour.

Available modes include:

- SOV
- HOV2
- HOV3+
- Auto Pay
- Transit
- Bike
- Walk

The model coefficients below are the result of multiple linear regression with a
forced intercept at zero. In addition, the model estimation will
be scaled up to a predicted regional total. As a result of these two factors,
the displayed r-squared values are not as meaningful. Instead, the value of the
coefficients is in determining the relative effect that various home-based trip
types have on non-home-based trip generation.

When reviewing the coefficients below, note their logical consistency: SOV NHB
trips are most likely to result when the HB trip is SOV or HOV. NHB walk trips
can be made when a person drives from home, but is more much more likely if they
walk. These results greatly improve the NHB models compared to traditional
trip-based construction.

```{r, include=FALSE}
trips_orig <- read_csv(
  "data/_private/hh_survey/output/output_trips.csv")
```


```{r, include = FALSE}
trips <- trips_orig %>%
  filter(!is.na(p_taz) & !is.na(a_taz) & purpose != "N_HBH") %>%
  mutate(
    mode = ifelse(mode %in% c("walk", "bike"), "walkbike", mode),
    mode = ifelse(mode == "tnc", "auto_pay", mode),
    purpose = paste0(purpose, "_", mode),
    trip_type = purpose,
    homebased = ifelse(grepl("_HB", trip_type), "HB", "NH")
  )

# hold all model tables and plots in a list
coeffs_list <- list()
p_tbl_list <- list()
```

```{r}
make_table <- function(tbl) {
  tbl %>%
    arrange(estimated_as) %>%
    kable(digits = 4) %>%
    kable_styling(full_width = FALSE)
}

# unique(trips$mode_nhb)
```

## Work tours

Each trip type below is a non-home-based trip on a work tour. Each tab shows a
mode-specific model that estimates which home-based trips and modes generate that
non-home-based trip and mode combination.

### W_NHBO {.tabset .W_NHBO}

These are non-home-based trips made on a work tour not related to work. In
other words, neither end of the trip is work or home.

#### SOV

```{r, W_NHBO_sov}
type <- "W_NHBO_sov"

# Collapse trip types
equiv <- list(
  "W_HBW_hov2" = "W_HBW_hov",
  "W_HBW_hov3" = "W_HBW_hov",
  "W_HBW_sov" = "W_HBW_sov",
  "W_HBO_hov2" = "W_HBO_hov",
  "W_HBO_hov3" = "W_HBO_hov"
)

# Restrict dependent variables
dependent_vars <- c(
  "W_HBO_sov",
  "W_HBW_sov"
)

# model1 <- estimate_nhb(trips, type)
# View(model1$tbl)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(trips, type, equiv = equiv, 
                       dependent_vars = dependent_vars)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```

#### HOV2

```{r, W_NHBO_hov2}
type <- "W_NHBO_hov2"

# Collapse trip types
equiv <- list(
  "W_HBO_hov2" = "W_HBO_auto",
  "W_HBO_hov3" = "W_HBO_auto",
  "W_HBO_sov" = "W_HBO_auto"
)

# Restrict dependent variables
dependent_vars <- c(
  "W_HBO_auto"
)

# model1 <- estimate_nhb(trips, type)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(trips, type, equiv = equiv, 
                       dependent_vars = dependent_vars)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```

#### HOV3+

```{r, W_NHBO_hov}
type <- "W_NHBO_hov3"

# Collapse trip types
equiv <- list(
  "W_HBO_hov2" = "W_HBO_auto",
  "W_HBO_hov3" = "W_HBO_auto",
  "W_HBO_sov" = "W_HBO_auto"
)

# Restrict dependent variables
dependent_vars <- c(
  "W_HBO_auto"
)

# model1 <- estimate_nhb(trips, type)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(trips, type, equiv = equiv, 
                       dependent_vars = dependent_vars)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```

#### Auto Pay

No significant presence of this trip type in the survey, and it is
not an important source of travel demand. No model was estimated.

```{r, W_NHBO_auto_pay}
type <- "W_NHBO_auto_pay"
n_records <- trips %>%
  filter(purpose == type) %>%
  nrow()
```

#### Transit

```{r, W_NHBO_bus}
type <- "W_NHBO_bus"

# Collapse trip types
equiv <- list(
  "W_HBO_hov2" = "W_HBO_auto",
  "W_HBO_hov3" = "W_HBO_auto",
  "W_HBO_sov" = "W_HBO_auto"
)

# Restrict dependent variables
dependent_vars <- c(
  "W_HBO_walkbike",
  "W_HBO_bus"
)

# model1 <- estimate_nhb(trips, type)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(trips, type, equiv = equiv, 
                       dependent_vars = dependent_vars)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```

#### Non-motorized

```{r, W_NH_EK12_All_walkbike}
type <- "W_NHBO_walkbike"

# Collapse trip types
equiv <- NULL

# Restrict dependent variables
dependent_vars <- c(
  "W_HBO_walkbike"
)

# model1 <- estimate_nhb(trips, type)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(trips, type, equiv = equiv, 
                       dependent_vars = dependent_vars)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```


### W_NHBW {.tabset .W_NHBW}

These are non-home-based trips made on a work tour where one trip end is related
to work, but not the primary work location. A trip made to attend a business
meeting is a good example.

#### SOV

```{r, W_NHBW_sov}
type <- "W_NHBW_sov"

# Collapse trip types
equiv <- list(
  "W_HBW_hov2" = "W_HBW_auto",
  "W_HBW_hov3" = "W_HBW_auto",
  "W_HBW_sov" = "W_HBW_auto"
)

# Restrict dependent variables
dependent_vars <- c(
  "W_HBO_sov",
  "W_HBO_hov2",
  "W_HBO_hov3",
  "W_HBW_auto"
)

# model1 <- estimate_nhb(trips, type)
# View(model1$tbl)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(trips, type, equiv = equiv, 
                       dependent_vars = dependent_vars)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```

#### HOV2

```{r, W_NHBW_hov2}
type <- "W_NHBW_hov2"

# Collapse trip types
equiv <- list(
  "W_HBO_hov2" = "W_HBO_hov",
  "W_HBO_hov3" = "W_HBO_hov"
)

# Restrict dependent variables
dependent_vars <- c(
  'W_HBO_hov',
  'W_HBO_sov'
)

# model1 <- estimate_nhb(trips, type)
# View(model1$tbl)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(trips, type, equiv = equiv, 
                       dependent_vars = dependent_vars)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```

#### HOV3+

```{r, W_NHBW_hov3}
type <- "W_NHBW_hov3"

# Collapse trip types
equiv <- list(
  "W_HBO_hov2" = "W_HBO_hov",
  "W_HBO_hov3" = "W_HBO_hov",
  "W_HBW_hov2" = "W_HBW_hov"
)

# Restrict dependent variables
dependent_vars <- c(
  'W_HBO_hov',
  "W_HBW_hov3"
)

# model1 <- estimate_nhb(trips, type)
# View(model1$tbl)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(trips, type, equiv = equiv, 
                       dependent_vars = dependent_vars)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```

#### Auto Pay

```{r, W_NHBW_auto_pay}
type <- "W_NHBW_auto_pay"

# Collapse trip types
equiv <- NULL

# Restrict dependent variables
dependent_vars <- c(
  'W_HBO_auto_pay'
)

# model1 <- estimate_nhb(trips, type)
# View(model1$tbl)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(trips, type, equiv = equiv, 
                       dependent_vars = dependent_vars)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```

#### Transit

```{r, W_NHBW_lb}
type <- "W_NHBW_bus"

# Collapse trip types
equiv <- list(
  "W_HBO_hov2" = "W_HBO_auto",
  "W_HBO_hov3" = "W_HBO_auto",
  "W_HBO_sov" = "W_HBO_auto"
)

# Restrict dependent variables
dependent_vars <- c(
  "W_HBO_bus"
)

# model1 <- estimate_nhb(trips, type)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(trips, type, equiv = equiv, 
                       dependent_vars = dependent_vars)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```

#### Non-motorized

```{r, W_NHBW_walkbike}
type <- "W_NHBW_walkbike"

# Collapse trip types
equiv <- list(
  "W_HBO_hov2" = "W_HBO_hov",
  "W_HBO_hov3" = "W_HBO_hov"
)

# Restrict dependent variables
dependent_vars <- c(
  "W_HBW_walkbike"
)

# model1 <- estimate_nhb(trips, type)
# View(model1$tbl)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(trips, type, equiv = equiv, 
                       dependent_vars = dependent_vars)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```

## Non-work tours

Each trip type below is a non-home-based trip made on a non-work tour. Each tab
shows a mode-specific model that estimates which home-based trips and modes
generate that non-home-based trip and mode combination.

For many of these trip types, the models have two additional parameters alpha
and beta. These coefficients control the impact of zonal accessibility on the
models and are discussed in detail in the next section.

### N_NHBSHP {.tabset .N_NHBSHP}

These are non-home-based trips made on a non-work tour with one trip end for
shopping.

#### SOV

```{r, N_NHBSHP_sov}
type <- "N_NHBSHP_sov"

# Collapse trip types
equiv <- list(
  "N_HBO_sov" = "N_HBO_sov",
  "N_HBO_hov2" = "N_HBO_hov",
  "N_HBO_hov3" = "N_HBO_hov",
  "N_HBSHP_hov2" = "N_HBSHP_auto",
  "N_HBSHP_hov3" = "N_HBSHP_auto",
  "N_HBSHP_sov" = "N_HBSHP_auto",
  "N_HBSR_hov2" = "N_HBSR_auto",
  "N_HBSR_hov3" = "N_HBSR_auto",
  "N_HBSR_sov" = "N_HBSR_auto"
)

# Restrict dependent variables
dependent_vars <- c(
  "N_HBO_sov",
  "N_HBSHP_auto",
  "N_HBSR_auto"
)

# model1 <- estimate_nhb(trips, type)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(
  trips, type, equiv = equiv, dependent_vars = dependent_vars,
  boost = TRUE
)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```



#### HOV2

```{r, N_NHBSHP_hov2}
type <- "N_NHBSHP_hov2"

# Collapse trip types
equiv <- list(
  "N_HBO_hov2" = "N_HBO_hov",
  "N_HBO_hov3" = "N_HBO_hov",
  "N_HBSHP_hov3" = "N_HBSHP_hov",
  "N_HBSHP_hov3" = "N_HBSHP_hov",
  "N_HBSR_hov3" = "N_HBSR_hov",
  "N_HBSR_hov3" = "N_HBSR_hov"
)

# Restrict dependent variables
dependent_vars <- c(
  "N_HBO_hov"
)

# model1 <- estimate_nhb(trips, type)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(
  trips, type, equiv = equiv, dependent_vars = dependent_vars,
  boost = TRUE
)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```

#### HOV3+

```{r, N_NHBSHP_hov3}
type <- "N_NHBSHP_hov3"

# Collapse trip types
equiv <- list(
  "N_HBSHP_hov3" = "N_HBSHP_hov",
  "N_HBSHP_hov3" = "N_HBSHP_hov"
)

# Restrict dependent variables
dependent_vars <- c(
  "N_HBSHP_hov",
  "N_HBO_hov2"
)

# model1 <- estimate_nhb(trips, type)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(
  trips, type, equiv = equiv, dependent_vars = dependent_vars,
  boost = TRUE
)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```

#### Auto Pay

```{r, N_NHBSHP_auto_pay}
type <- "N_NHBSHP_auto_pay"

# Collapse trip types
equiv <- NULL

# Restrict dependent variables
dependent_vars <- c(
  "N_HBSHP_auto_pay"
)

# model1 <- estimate_nhb(trips, type)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(
  trips, type, equiv = equiv, dependent_vars = dependent_vars,
  boost = TRUE
)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```

#### Transit

```{r, N_NHBSHP_lb}
type <- "N_NHBSHP_bus"


# Collapse trip types
equiv <- NULL

# Restrict dependent variables
dependent_vars <- c(
  "N_HBSHP_bus"
)

# model1 <- estimate_nhb(trips, type)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(
  trips, type, equiv = equiv, dependent_vars = dependent_vars,
  boost = FALSE
)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```

#### Non-motorized

```{r, N_NHBSHP_walkbike}
type <- "N_NHBSHP_walkbike"

# Collapse trip types
equiv <- NULL

# Restrict dependent variables
dependent_vars <- c(
  "N_HBSHP_walkbike"
)

# model1 <- estimate_nhb(trips, type)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(
  trips, type, equiv = equiv, dependent_vars = dependent_vars,
  boost = FALSE
)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```


### N_NHBO {.tabset .N_NHBO}

These non-home-based trips are those where neither end is shopping. In effect,
these are all other non-home-based trips made on non-work tours. Visiting
friends or family at their home is an example of this trip type.

#### SOV

```{r, N_NHBO_sov}
type <- "N_NHBO_sov"

# Collapse trip types
equiv <- c(
  "N_HBO_hov2" = "N_HBO_auto",
  "N_HBO_hov3" = "N_HBO_auto",
  "N_HBO_sov" = "N_HBO_auto",
  "N_HBSR_hov2" = "N_HBSR_auto",
  "N_HBSR_hov3" = "N_HBSR_auto",
  "N_HBSR_sov" = "N_HBSR_auto",
  "N_HBSHP_hov2" = "N_HBSHP_auto",
  "N_HBSHP_hov3" = "N_HBSHP_auto",
  "N_HBSHP_sov" = "N_HBSHP_auto"
)

# Restrict dependent variables
dependent_vars <- c(
  "N_HBO_auto",
  "N_HBSR_auto"
)

# model1 <- estimate_nhb(trips, type)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(
  trips, type, equiv = equiv, dependent_vars = dependent_vars,
  boost = TRUE
)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```

#### HOV2

```{r, N_NHBO_hov2}
type <- "N_NHBO_hov2"

# Collapse trip types
equiv <- c(
  "N_HBO_hov2" = "N_HBO_auto",
  "N_HBO_hov3" = "N_HBO_auto",
  "N_HBO_sov" = "N_HBO_auto",
  "N_HBSR_hov2" = "N_HBSR_hov",
  "N_HBSR_hov3" = "N_HBSR_hov",
  "N_HBO_hov2" = "N_HBO_hov",
  "N_HBO_hov3" = "N_HBO_hov",
  "N_HBSHP_hov2" = "N_HBSHP_hov",
  "N_HBSHP_hov3" = "N_HBSHP_hov"
)

# Restrict dependent variables
dependent_vars <- c(
  "N_HBO_auto",
  "N_HBSR_hov",
  "N_HBSHP_sov"
)

# model1 <- estimate_nhb(trips, type)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(
  trips, type, equiv = equiv, dependent_vars = dependent_vars,
  boost = TRUE
)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```

#### HOV3+

```{r, N_NHBO_hov3}
type <- "N_NHBO_hov3"

# Collapse trip types
equiv <- c(
  "N_HBO_hov2" = "N_HBO_auto",
  "N_HBO_hov3" = "N_HBO_auto",
  "N_HBO_sov" = "N_HBO_auto",
  "N_HBSHP_sov" = "N_HBSHP_auto",
  "N_HBSHP_hov2" = "N_HBSHP_auto",
  "N_HBSHP_hov3" = "N_HBSHP_auto"
)

# Restrict dependent variables
dependent_vars <- c(
  "N_HBO_auto",
  "N_HBSHP_auto"
)

# model1 <- estimate_nhb(trips, type)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(
  trips, type, equiv = equiv, dependent_vars = dependent_vars,
  boost = TRUE
)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```


#### Auto Pay

```{r, N_NHBO_auto_pay}
type <- "N_NHBO_auto_pay"

# Collapse trip types
equiv <- NULL

# Restrict dependent variables
dependent_vars <- c(
  "N_HBO_auto_pay",
  "N_HBSHP_auto_pay",
  "N_HBSR_auto_pay"
)

# model1 <- estimate_nhb(trips, type)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(
  trips, type, equiv = equiv, dependent_vars = dependent_vars,
  boost = FALSE
)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```

#### Transit

```{r, N_NHBO_lb}
type <- "N_NHBO_bus"

# Collapse trip types
equiv <- NULL

# Restrict dependent variables
dependent_vars <- c(
  "N_HBO_bus",
  "N_HBSR_bus",
  "N_HBO_walkbike"
)

# model1 <- estimate_nhb(trips, type)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(
  trips, type, equiv = equiv, dependent_vars = dependent_vars,
  boost = FALSE
)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[gsub("_bus", "_t", type)] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```



#### Non-motorized

```{r, N_NHBO_walkbike}
type <- "N_NHBO_walkbike"

# Collapse trip types
equiv <- NULL

# Restrict dependent variables
dependent_vars <- c(
  "N_HBO_walkbike"
)

# model1 <- estimate_nhb(trips, type)
# model2 <- estimate_nhb(trips, type, equiv = equiv)
model3 <- estimate_nhb(
  trips, type, equiv = equiv, dependent_vars = dependent_vars,
  boost = FALSE
)

zonal_result <- apply_nhb_zonal(trips, type, model3$tbl)
coeff_tbl <- model3$tbl %>%
  add_row(term = "r_sq", estimate = zonal_result$r_sq)
coeffs_list[type] <- list(coeff_tbl)
make_table(coeff_tbl)
p_tbl_list[type] <- list(model3$p_tbl)
```

## Boosting charts

"Boosting" is an approach borrowed from machine learning where the errors of a
previous model are used to estimate a second model. The models shown above
illustrate the link between home- and non-home-based trips by mode, but
intuition and experience tells us that accessibility should also influence NHB
trip making. Travel to a central business district is more likely to lead to
further trips compared to traveling to a rural zone.

The chart below shows how accessibility impacts each trip type and mode
combination. The y-axis is a simple factor. When the factor is 1, the trip rates
will be the same as displayed in the tables above. A y-value of 0.5 means the
trip rates will be reduced by 50 percent. Conversely, a y-value of 1.5 means
that trip rates are increased by 50 percent. With this additional factor, the
model will understand the role that accessibility plays in NHB trip generation.

```{r}
tbl <- bind_rows(p_tbl_list)
ggplot(tbl) +
  geom_point(aes(x = logsum, y = y, color = trip_type)) +
  labs(
    title = "NHB Trip-Making and Accessibility",
    y = "Effect on NHB Trip Rate",
    x = "Nearby Accessibility",
    color = "Trip Type"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
```

## Time of Day Adjustment

The independent feedback by time of day in the model presents a unique challenge
for NHB. NHB and HB trips have very different time of day patterns (see Time of
Day documentation), but NHB generation is simply rates multiplied by HB trip
ends. Without correction, this would mean that the NHB trips would have the same
distribution as HB trips. To correct this, Caliper calculated adjustment factors
by tour type, mode, and time of day. This was done by comparing raw model
outputs back to the observed NHB trips from the survey.

The easiest way to understand the table below is by highlighting one row as an
example. For SOV trips on work tours, these factors move NHB trips out of the AM
and NT periods and into the MD and PM. Another way of saying it is that,
compared to HB trips, NHB trips are less likely in the AM/NT and more likely in
the MD/PM. The large mid day factor reflects lunch trips and other mid-day
activities. The relative size of the AM and PM periods confirms what we know:
that stops are more likely to be made on the way home compared to the way to
work (e.g. picking up groceries).

```{r, include=FALSE}
# This file is produced by the model. Importantly, it must be generated
# by a model run with calibration factors set to 1.
results <- read_csv("data/input/nhb/nhb_generation.csv") %>%
  select(-c(TAZ:access_walk))
```


```{r}
model <- results %>%
  pivot_longer(names_to = "type", cols = everything()) %>%
  mutate(type = gsub("auto_pay", "autopay", type)) %>%
  separate(type, into = c("tour_type", "hb", "activity", "length", "mode", "tod"), remove = FALSE) %>%
  group_by(tour_type, mode, tod) %>%
  summarize(est_trips = sum(value))

comparison <- trips %>%
  mutate(
    # mode_nhb = mode
    mode_nhb = case_when(
      homebased == "HB" ~ mode,
      mode == "auto_pay" ~ "autopay",
      mode != "other_auto" ~ mode,
      party_size == 1 ~ "sov",
      party_size == 2 ~ "hov2",
      party_size >= 3 ~ "hov3"
    )
  ) %>%
  filter(
    homebased == "NH",
    tour_type != "H",
    !(mode_nhb %in% c("school_bus", "XXX", "other"))
  ) %>%
  mutate(
    mode_nhb = ifelse(mode_nhb %in% c("lb", "eb"), "t", mode_nhb)
  ) %>%
  group_by(tour_type, tod, mode_nhb) %>%
  summarize(obs_trips = sum(trip_weight_combined)) %>%
  rename(mode = mode_nhb) %>%
  left_join(
    model, 
    by = c("tour_type" = "tour_type", "tod" = "tod", "mode" = "mode")
  ) %>%
  mutate(
    diff = est_trips - obs_trips,
    pct_diff = (diff / obs_trips) * 100,
    factor = round(obs_trips / est_trips, 4)
  ) %>%
  ungroup()

comparison %>%
  select(-diff, -pct_diff) %>%
  mutate(
    obs_trips = round(obs_trips, 0),
    est_trips = round(est_trips, 0),
    mode = case_when(
      mode == "autopay" ~ "Auto Pay",
      mode == "sov" ~ "SOV",
      mode == "hov2" ~ "HOV2",
      mode == "hov3" ~ "HOV3",
      mode == "t" ~ "Transit",
      mode == "walkbike" ~ "Nonmotorized",
    ),
    mode = factor(
      mode, 
      levels = c("SOV", "HOV2", "HOV3", "Auto Pay", "Transit", "Nonmotorized"),
      ordered = TRUE
    )
  ) %>%
  select(tour_type, tod, mode, factor) %>%
  pivot_wider(names_from = tod, values_from = factor) %>%
  mutate(tour_type = ifelse(tour_type == "N", "NonWork", "Work")) %>%
  rename(
    Mode = mode,
    `Tour Type` = tour_type
  ) %>%
  kable(digits = 2) %>%
  kable_styling(full_width = FALSE)

calib_out_tbl <- comparison %>%
  mutate(
    mode = ifelse(mode == "autopay", "auto_pay", mode),
    type = paste(tour_type, mode, tod, sep = "_")
  ) %>%
  select(type, factor)
```

```{r write_output, eval=FALSE}
output_dir <- "../master/resident/nhb/generation/"
for (i in 1:length(coeffs_list)) {
  name <- names(coeffs_list)[i]
  df <- coeffs_list[[i]]
  write_csv(df, paste0(output_dir, name, ".csv"))
}

write_csv(calib_out_tbl, paste0(output_dir, "nhb_calibration_factors.csv"))
```