---
title: "Reno Production Trees"
output: html_document
date: "2024-08-22"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(dplyr.summarise.inform = FALSE)
options(scipen = 999)

library(tidyverse)
library(rpart)
library(rpart.plot)
library(gsubfn)
library(kableExtra)
```

```{r, include=FALSE}
hh <- read_csv("data/_private/hh_survey/output/output_households.csv")
per <- read_csv("data/_private/hh_survey/output/output_persons.csv") %>%
  mutate(
    person_id = paste0(Sample_Number, "_", Person_Number),
    is_kid = ifelse(Age < 18, 1, 0),
    is_senior = ifelse(Age >= 65, 1, 0),
    is_adult = ifelse(Age >= 18, 1, 0)
  )
trips <- read_csv("data/_private/hh_survey/output/output_trips.csv") %>%
  mutate(person_id = paste0(Sample_Number, "_", Person_Number))
accessibily <- read_csv("data/input/accessibility/accessibilities.csv")
```

```{r}
# Aggregate counts of trips to persons
agg_trips <- trips %>%
  group_by(person_id, purpose) %>%
  summarize(trips = n()) %>%
  ungroup() %>%
  pivot_wider(id_cols = person_id, names_from = purpose, values_from = trips, values_fill = 0)

# Determine which households have kids
has_kids <- per %>%
  group_by(Sample_Number) %>%
  summarize(
    num_kids = sum(is_kid, na.rm = TRUE),
    num_seniors = sum(is_senior, na.rm = TRUE),
    num_adults = sum(is_adult, na.rm = TRUE)
  ) %>%
  mutate(
    kids_in_hh = ifelse(num_kids > 0, "Yes", "No"),
    seniors_in_hh = ifelse(num_seniors > 0, "Yes", "No")
  )
```


```{r}
# join the three sets of survey info
joined <- per %>%
  left_join(hh, by = "Sample_Number") %>%
  left_join(has_kids, by = "Sample_Number") %>%
  # set factor fields
  mutate(
    Gender = ifelse(Gender == 1, "Male", "Female"),
    Gender = as.factor(Gender),
    Student_Status = as.factor(Student_Status),
    Employed = ifelse(Employment == "Yes", "Yes", "No"),
    Student = Student_Status,
    Regular_WorkAtHome = ifelse(Workplace_Type_int == 1, "Yes", "No"),
    WorkIndustry = as.factor(Primary_Industry_int),
    HHIncCat = case_when(
      Income_int.x <= 4 ~ 1,
      Income_int.x <= 9 ~ 2,
      Income_int.x <= 14 ~ 3,
      Income_int.x <= 20 ~ 4,
      Income_int.x <= 24 ~ 5,
      TRUE ~ 6
    ),
    IncPerCapita = income_midpt.x / Number_Persons.x,
    OwnOrRent = as.factor(Rent_or_Own_Residence.x),
    VehSufficiency = case_when(
      `Vehicles_Available.x` == 0 ~ "v0",
      `Vehicles_Available.x` < num_adults ~ "vi",
      TRUE ~ "vs"
    )
  ) %>%
  left_join(agg_trips, by = "person_id") %>%
  mutate(across(W_HBW:N_HBH, ~ replace_na(.x, 0))) %>%
  left_join(accessibily, by = c("HomeTAZ.x" = "TAZ")) %>%
  mutate(
    is_kid = ifelse(is_kid == 1, "Yes", "No"),
    is_adult = ifelse(is_adult == 1, "Yes", "No"),
    is_senior = ifelse(is_senior == 1, "Yes", "No"),
  ) %>%
  select(
    person_id, 
    # HH variables
    HHSize = Number_Persons.x, 
    Vehicles = `Vehicles_Owned/Leased.x`, 
    HHIncCat,
    IncPerCapita,
    HHKids = num_kids,
    HHSeniors = num_seniors,
    HHAdults = num_adults,
    kids_in_hh,
    seniors_in_hh,
    # person variables
    Gender,
    Age,
    is_kid,
    is_adult,
    is_senior,
    Employed,
    g_access = access_general_sov,
    t_access = access_transit,
    w_access = access_walk,
    VehSufficiency,
    W_HBW:N_HBH,
    WeightCaliper = `Weight_Factor.x`
  ) %>%
  filter(!is.na(WeightCaliper))
```

```{r}
# Extract the terminal nodes and their average rates
create_rate_table <- function(fit, market_segment = NULL) {
  frame <- fit$frame
  leaves <- which(frame$var == "<leaf>")
  
  # Initialize an empty data frame to store the results
  results <- data.frame(rule = character(), avg_rate = numeric(), stringsAsFactors = FALSE)
  
  for (leaf in leaves) {
    path <- path.rpart(fit, nodes = rownames(frame)[leaf], print.it = FALSE)
    
    # Function to convert scientific notation in rule conditions to normal numbers
    clean_rule <- function(rule) {
      gsubfn(pattern = "(\\d+\\.\\d+e[+-]\\d+)", 
           replacement = function(x) format(as.numeric(x), scientific = FALSE), 
           rule)
    }
    
    path <- unlist(path)
    path <- path[2:length(path)]
    rule <- paste(sapply(path, clean_rule), collapse = " and ")
    
    rate <- frame$yval[leaf]
    results <- rbind(results, data.frame(rule = rule, rate = rate))
  }
  results$rate <- round(results$rate, 4)
  
  if (!is.null(market_segment)) {
    results <- results %>%
      mutate(rule = paste0("auto_suff = '", market_segment, "' and ", rule))
  }
  
  return(results)
}

rate_list <- list()
```

The following analysis shows the affects of estimated decision trees on different
segments of the population. In general, pre-segmentation of the population
improves model fit when comparing back to the survey by market segment.

# W_HBW by Vehicle Sufficiency {.tabset}

## v0

```{r}
est_tbl <- joined %>%
  select(
    W_HBW, WeightCaliper,
    HHSize:VehSufficiency
  ) %>%
  filter(VehSufficiency == "v0")
  

tree <- rpart(
  W_HBW ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 5,
  cp = .001,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, 2)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rate_tbl <- create_rate_table(tree, "v0") %>%
  mutate(trip_type = "W_HBW_v0") %>%
  relocate(trip_type, .before = rule)
rate_list$W_HBW_v0 <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

## vi

```{r}
est_tbl <- joined %>%
  select(
    W_HBW, WeightCaliper,
    HHSize:VehSufficiency,
    -w_access, -HHSize
  ) %>%
  filter(VehSufficiency == "vi")

tree <- rpart(
  W_HBW ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 5,
  cp = .001,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, 6)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rate_tbl <- create_rate_table(tree, "vi") %>%
  mutate(trip_type = "W_HBW_vi") %>%
  relocate(trip_type, .before = rule)
rate_list$W_HBW_vi <- rate_tbl
```

## vs

```{r}
est_tbl <- joined %>%
  select(
    W_HBW, WeightCaliper,
    HHSize:VehSufficiency
  ) %>%
  mutate(over_60 = ifelse(Age >= 61, "Yes", "No")) %>%
  select(-Age) %>%
  filter(VehSufficiency == "vs")

tree <- rpart(
  W_HBW ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 5,
  cp = .001,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, c(29, 31))
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rate_tbl <- create_rate_table(tree, "vs") %>%
  mutate(trip_type = "W_HBW_vs") %>%
  relocate(trip_type, .before = rule)
rate_list$W_HBW_vs <- rate_tbl
```

# W_HBO by Vehicle Sufficiency {.tabset}

## v0

```{r}
est_tbl <- joined %>%
  select(
    W_HBO, WeightCaliper,
    HHSize:VehSufficiency
  ) %>%
  filter(VehSufficiency == "v0")
  

tree <- rpart(
  W_HBO ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 2,
  cp = .001,
  minbucket = 30,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
# tree <- snip.rpart(tree, 2)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rate_tbl <- create_rate_table(tree, "v0") %>%
  mutate(trip_type = "W_HBO_v0") %>%
  relocate(trip_type, .before = rule)
rate_list$W_HBO_v0 <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

## vi

```{r}
est_tbl <- joined %>%
  select(
    W_HBO, WeightCaliper,
    HHSize:VehSufficiency,
    -w_access
  ) %>%
  filter(VehSufficiency == "vi")
  

tree <- rpart(
  W_HBO ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 5,
  cp = .01,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, 29)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rate_tbl <- create_rate_table(tree, "vi") %>%
  mutate(trip_type = "W_HBO_vi") %>%
  relocate(trip_type, .before = rule)
rate_list$W_HBO_vi <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

## vs

```{r}
est_tbl <- joined %>%
  select(
    W_HBO, WeightCaliper,
    HHSize:VehSufficiency
  ) %>%
  filter(VehSufficiency == "vs")
  

tree <- rpart(
  W_HBO ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 4,
  cp = .01,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
# tree <- snip.rpart(tree, 2)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rate_tbl <- create_rate_table(tree, "vs") %>%
  mutate(trip_type = "W_HBO_vs") %>%
  relocate(trip_type, .before = rule)
rate_list$W_HBO_vs <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

# N_HBSR by Vehicle Sufficiency {.tabset}

## v0

```{r}
est_tbl <- joined %>%
  select(
    N_HBSR, WeightCaliper,
    HHSize:VehSufficiency
  ) %>%
  filter(VehSufficiency == "v0")
  

tree <- rpart(
  N_HBSR ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 5,
  cp = .02,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, 6)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rate_tbl <- create_rate_table(tree, "v0") %>%
  mutate(trip_type = "N_HBSR_v0") %>%
  relocate(trip_type, .before = rule)
rate_list$N_HBSR_v0 <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

## vi

```{r}
est_tbl <- joined %>%
  select(
    N_HBSR, WeightCaliper,
    HHSize:VehSufficiency,
    -t_access
  ) %>%
  filter(VehSufficiency == "vi")
  

tree <- rpart(
  N_HBSR ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 4,
  cp = .01,
  minbucket = 59,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, c(6, 14))
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rate_tbl <- create_rate_table(tree, "vi") %>%
  mutate(trip_type = "N_HBSR_vi") %>%
  relocate(trip_type, .before = rule)
rate_list$N_HBSR_vi <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

## vs

```{r}
est_tbl <- joined %>%
  mutate(IncPerCapLt80 = ifelse(IncPerCapita < 80000, "Yes", "No")) %>%
  select(
    N_HBSR, WeightCaliper,
    HHSize:VehSufficiency,
    -HHSize, -Age, -HHIncCat, -IncPerCapita,
    IncPerCapLt80,
    -Vehicles, -t_access, -g_access
  ) %>%
  filter(VehSufficiency == "vs")
  

tree <- rpart(
  N_HBSR ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 4,
  cp = .007,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, 13)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rate_tbl <- create_rate_table(tree, "vs") %>%
  mutate(trip_type = "N_HBSR_vs") %>%
  relocate(trip_type, .before = rule)
rate_list$N_HBSR_vs <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

# N_HBO by Vehicle Sufficiency {.tabset}

## v0

```{r}
est_tbl <- joined %>%
  select(
    N_HBO, WeightCaliper,
    HHSize:VehSufficiency
    # -w_access
  ) %>%
  mutate(t_access_lt_61 = ifelse(t_access < 6.1, "Yes", "No")) %>%
  select(-t_access, -w_access) %>%
  filter(VehSufficiency == "v0")
  

tree <- rpart(
  N_HBO ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 4,
  cp = .01,
  minbucket = 30,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
# tree <- snip.rpart(tree, 2)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rate_tbl <- create_rate_table(tree, "v0") %>%
  mutate(trip_type = "N_HBO_v0") %>%
  relocate(trip_type, .before = rule)
rate_list$N_HBO_v0 <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

## vi

```{r}
est_tbl <- joined %>%
  select(
    N_HBO, WeightCaliper,
    HHSize:VehSufficiency,
    -IncPerCapita, -HHIncCat
  ) %>%
  filter(VehSufficiency == "vi")
  

tree <- rpart(
  N_HBO ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 4,
  cp = .01,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
# tree <- snip.rpart(tree, 2)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rate_tbl <- create_rate_table(tree, "vi") %>%
  mutate(trip_type = "N_HBO_vi") %>%
  relocate(trip_type, .before = rule)
rate_list$N_HBO_vi <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

## vs

```{r}
est_tbl <- joined %>%
  select(
    N_HBO, WeightCaliper,
    HHSize:VehSufficiency
  ) %>%
  filter(VehSufficiency == "vs")
  

tree <- rpart(
  N_HBO ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 4,
  cp = .01,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
# tree <- snip.rpart(tree, 2)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rate_tbl <- create_rate_table(tree, "vs") %>%
  mutate(trip_type = "N_HBO_vs") %>%
  relocate(trip_type, .before = rule)
rate_list$N_HBO_vs <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

# N_HBSCH

```{r}
est_tbl <- joined %>%
  select(
    N_HBSCH, WeightCaliper,
    HHSize:VehSufficiency, -HHSize, -HHKids, -t_access
  )

tree <- rpart(
  N_HBSCH ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 4,
  cp = .001,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, 15)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rate_tbl <- create_rate_table(tree) %>%
  mutate(trip_type = "N_HBSCH") %>%
  relocate(trip_type, .before = rule)
rate_list$N_HBSCH <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

# N_HBSHP by Vehicle Sufficiency {.tabset}

## v0

```{r}
est_tbl <- joined %>%
  select(
    N_HBSHP, WeightCaliper,
    HHSize:VehSufficiency, -IncPerCapita, -w_access, -t_access
  ) %>%
  filter(VehSufficiency == "v0")
  

tree <- rpart(
  N_HBSHP ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 5,
  cp = .001,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, 12)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rate_tbl <- create_rate_table(tree, "v0") %>%
  mutate(trip_type = "N_HBSHP_v0") %>%
  relocate(trip_type, .before = rule)
rate_list$N_HBSHP_v0 <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

## vi

```{r}
est_tbl <- joined %>%
  select(
    N_HBSHP, WeightCaliper,
    HHSize:VehSufficiency, -IncPerCapita,
    -HHAdults, -contains("access")
  ) %>%
  filter(VehSufficiency == "vi")
  

tree <- rpart(
  N_HBSHP ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 4,
  cp = .001,
  minbucket = 30,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, c(9, 10, 11))
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rate_tbl <- create_rate_table(tree, "vi") %>%
  mutate(trip_type = "N_HBSHP_vi") %>%
  relocate(trip_type, .before = rule)
rate_list$N_HBSHP_vi <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

## vs

```{r}
est_tbl <- joined %>%
  select(
    N_HBSHP, WeightCaliper,
    HHSize:VehSufficiency,
    -Vehicles, -w_access, -IncPerCapita, -t_access
  ) %>%
  filter(VehSufficiency == "vs")
  

tree <- rpart(
  N_HBSHP ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 5,
  cp = .007,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, c(9, 5))
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rate_tbl <- create_rate_table(tree, "vs") %>%
  mutate(trip_type = "N_HBSHP_vs") %>%
  relocate(trip_type, .before = rule)
rate_list$N_HBSHP_vs <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

# Calibration

Calibrating the production rates is done to ensure that the final model is
producing the same number of trips per person on average as the survey. One
complicating factor in this comparison is that the survey and model
socio-economic data have different total populations.

- Survey: 476,187
- Model: 523,542

```{r}
survey_pop <- sum(per$Weight_Factor)
model_pop <- 523542
pop_factor <- model_pop / survey_pop
```

As a consequence, the survey trip totals are increased by `r round(pop_factor,
2)` and then compared to model results. The table below shows the production
model results compared back to the total trips in the survey (factored up).
While some purposes are higher or lower than the survey, the model does a good
job of predicing total trips even before calibration.

```{r, include=FALSE}
# Importantly, this file must be from a model run with calibration factors
# set to 1 (no adjustment).
model_trips <- read_csv(
  "data/input/production_trees/modeled_trips_uncalibrated.csv") %>%
  pivot_longer(
    N_HBO_v0:W_HBW_vs, names_to = "trip_type", values_to = "model") %>%
  group_by(trip_type) %>%
  summarize(model = sum(model, na.rm = TRUE)) %>%
  mutate(
    segment = str_sub(trip_type, -2),
    segment = ifelse(trip_type == "N_HBSCH", "", segment),
    trip_type = case_when(
      trip_type == "N_HBSCH" ~ trip_type,
      TRUE ~ str_sub(trip_type, 1, -4)
    )
  )
```

```{r}
temp <- hh %>%
  left_join(has_kids, by = "Sample_Number") %>%
  mutate(VehSufficiency = case_when(
    `Vehicles_Available` == 0 ~ "v0",
    `Vehicles_Available` < num_adults ~ "vi",
    TRUE ~ "vs"
  )) %>%
  select(Sample_Number, VehSufficiency)

survey_trips <- trips %>%
  filter(purpose != "N_HBH" & !grepl("NHB", purpose)) %>%
  left_join(temp, by = "Sample_Number") %>%
  mutate(VehSufficiency = ifelse(purpose == "N_HBSCH", "", VehSufficiency)) %>%
  group_by(purpose, segment = VehSufficiency) %>%
  summarize(survey = sum(Weight_Factor) * pop_factor)

calibration_trips <- survey_trips %>%
  rename(trip_type = purpose) %>%
  left_join(model_trips, by = c("trip_type", "segment")) %>%
  mutate(
    diff = model - survey,
    `pct_diff` = round(diff / survey * 100, 1)
  ) %>%
  ungroup()

rollup <- calibration_trips %>%
  group_by(trip_type) %>%
  summarize(
    survey = sum(survey),
    model = sum(model)
  ) %>%
  mutate(
    diff = model - survey,
    `pct_diff` = round(diff / survey * 100, 1)
  ) %>%
  ungroup()

total <- rollup %>%
  summarize(
    trip_type = "All",
    survey = sum(survey),
    model = sum(model)
  ) %>%
  mutate(
    diff = model - survey,
    `pct_diff` = round(diff / survey * 100, 1)
  )

total <- bind_rows(rollup, total)

total %>%
  rename(
    `Trip Type` = trip_type,
    Observed = survey,
    Modelled = model,
    Difference = diff,
    `%Difference` = pct_diff
  ) %>%
  kable(digits = 0, format.args = list(big.mark = ",")) %>%
  kable_styling(full_width = FALSE)
```

The calibration factors below are the ratio of the observed to modeled trips.
These are applied by trip type and auto sufficiency segment to ensure total trip
making matches the survey (based on trip weight).

```{r}
calib_factors <- calibration_trips %>%
  mutate(factor = survey / model) %>%
  select(trip_type, segment, factor) %>%
  mutate(
    # factor = factor * 0.963, # final validation factor based on counts
    factor = round(factor, 3)
  ) 

calib_factors %>%
  rename(`Trip Type` = trip_type, Factor = factor) %>%
  kable(digits = 2) %>%
  kable_styling(full_width = FALSE)
```

```{r, eval=FALSE}
out_df <- bind_rows(rate_list) %>%
  mutate(
    rule = gsub("=No", "=0", rule, fixed = TRUE),
    rule = gsub("=Yes", "=1", rule, fixed = TRUE),
    rule = gsub("=Male", "=1", rule, fixed = TRUE),
    rule = gsub("=Female", "=2", rule, fixed = TRUE),
  )
write_csv(out_df, "data/output/production_trees/production_rates.csv")

write_csv(
  calib_factors,
  "data/output/production_trees/calibration_factors.csv"
)
```