---
title: "Reno Production Trees"
output: html_document
date: "2024-08-22"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(dplyr.summarise.inform = FALSE)
options(scipen = 999)

library(tidyverse)
library(rpart)
library(rpart.plot)
```

```{r, include=FALSE}
hh <- read_csv("data/_private/hh_survey/output/output_households.csv")
per <- read_csv("data/_private/hh_survey/output/output_persons.csv") %>%
  mutate(
    person_id = paste0(Sample_Number, "_", Person_Number),
    is_kid = ifelse(Age < 18, 1, 0),
    is_senior = ifelse(Age >= 65, 1, 0),
    is_adult = ifelse(Age >= 18, 1, 0)
  )
trips <- read_csv("data/_private/hh_survey/output/output_trips.csv") %>%
  mutate(person_id = paste0(Sample_Number, "_", Person_Number))
accessibily <- read_csv("data/input/accessibility/accessibilities.csv")
```

```{r}
# Aggregate counts of trips to persons
agg_trips <- trips %>%
  group_by(person_id, purpose) %>%
  summarize(trips = n()) %>%
  ungroup() %>%
  pivot_wider(id_cols = person_id, names_from = purpose, values_from = trips, values_fill = 0)

# Determine which households have kids
has_kids <- per %>%
  group_by(Sample_Number) %>%
  summarize(
    num_kids = sum(is_kid, na.rm = TRUE),
    num_seniors = sum(is_senior, na.rm = TRUE),
    num_adults = sum(is_adult, na.rm = TRUE)
  ) %>%
  mutate(
    kids_in_hh = ifelse(num_kids > 0, "Yes", "No"),
    seniors_in_hh = ifelse(num_seniors > 0, "Yes", "No")
  )
```


```{r}
# join the three sets of survey info
joined <- per %>%
  left_join(hh, by = "Sample_Number") %>%
  left_join(has_kids, by = "Sample_Number") %>%
  # set factor fields
  mutate(
    Gender = ifelse(Gender == 1, "Male", "Female"),
    Gender = as.factor(Gender),
    Student_Status = as.factor(Student_Status),
    Employed = ifelse(Employment == "Yes", "Yes", "No"),
    Student = Student_Status,
    Regular_WorkAtHome = ifelse(Workplace_Type_int == 1, "Yes", "No"),
    WorkIndustry = as.factor(Primary_Industry_int),
    HHIncCat = case_when(
      Income_int.x <= 4 ~ 1,
      Income_int.x <= 9 ~ 2,
      Income_int.x <= 14 ~ 3,
      Income_int.x <= 20 ~ 4,
      Income_int.x <= 24 ~ 5,
      TRUE ~ 6
    ),
    IncPerCapita = income_midpt.x / Number_Persons.x,
    OwnOrRent = as.factor(Rent_or_Own_Residence.x),
    # VehSufficiency = case_when(
    #   `Vehicles_Owned/Leased.x` == 0 ~ "v0",
    #   `Vehicles_Owned/Leased.x` < num_adults ~ "vi",
    #   TRUE ~ "vs"
    # )
    VehSufficiency = case_when(
      `Vehicles_Available.x` == 0 ~ "v0",
      `Vehicles_Available.x` < num_adults ~ "vi",
      TRUE ~ "vs"
    )
  ) %>%
  left_join(agg_trips, by = "person_id") %>%
  mutate(across(W_HBW:N_HBH, ~ replace_na(.x, 0))) %>%
  left_join(accessibily, by = c("HomeTAZ.x" = "TAZ")) %>%
  mutate(
    is_kid = ifelse(is_kid == 1, "Yes", "No"),
    is_adult = ifelse(is_adult == 1, "Yes", "No"),
    is_senior = ifelse(is_senior == 1, "Yes", "No"),
  ) %>%
  select(
    person_id, 
    # HH variables
    HHSize = Number_Persons.x, 
    Vehicles = `Vehicles_Owned/Leased.x`, 
    HHIncCat,
    IncPerCapita,
    HHKids = num_kids,
    HHSeniors = num_seniors,
    HHAdults = num_adults,
    kids_in_hh,
    seniors_in_hh,
    # person variables
    Gender,
    Age,
    is_kid,
    is_adult,
    is_senior,
    Employed,
    g_access = access_general_sov,
    t_access = access_transit,
    w_access = access_walk,
    VehSufficiency,
    W_HBW:N_HBH,
    WeightCaliper = `Weight_Factor.x`
  ) %>%
  filter(!is.na(WeightCaliper))
```

```{r}
rate_list <- list()

create_rule_table <- function(tree, market_segment = NULL) {
  rules <- rpart.rules(tree, digits = 4)
  colnames(rules) <- c(
    "rate",
    paste0("c", seq(2:ncol(rules) - 1))
  )

  
  result <- rules %>%
    as_tibble() %>%
    unite(col = "rule", -rate, sep = " ", na.rm = TRUE) %>%
    mutate(
      rate = round(as.numeric(rate), 4),
      rule = gsub("\\s+", " ", rule),
      rule = gsub("&", "and", rule),
      rule = gsub("when ", "", rule, fixed = TRUE),
      rule = gsub(" or ", ",", rule, fixed = TRUE),
      rule = str_replace(rule, "(\\w+)\\sis\\s(\\d+\\.*\\d*)\\sto\\s(\\d+\\.*\\d*)", "\\1 >= \\2 and \\1 < \\3"),
      rule = gsub(" is ", " = ", rule, fixed = TRUE),
      rate = as.numeric(rate)
    )
  
  if (!is.null(market_segment)) {
    result <- result %>%
      mutate(rule = paste0("auto_suff = '", market_segment, "' and ", rule))
  }
  
  result <- result %>%
    mutate(category = sort(unique(tree$where)))
  
  return(result)
}
```

The following analysis shows the affects of estimated decision trees on different
segments of the population. In general, pre-segmentation of the population
improves model fit when comparing back to the survey by market segment.

# W_HBW by Vehicle Sufficiency {.tabset}

## v0

```{r}
est_tbl <- joined %>%
  select(
    W_HBW, WeightCaliper,
    HHSize:VehSufficiency
  ) %>%
  filter(VehSufficiency == "v0")
  

tree <- rpart(
  W_HBW ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 5,
  cp = .001,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, 2)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rule_tbl <- create_rule_table(tree, "v0") %>%
  select(-rate)
rate_tbl <- est_tbl %>%
  mutate(category = tree$where,) %>%
  group_by(category) %>%
  dplyr::summarize(
    rate = round(weighted.mean(W_HBW, WeightCaliper), 4),
    stdev = round(sd(W_HBW, na.rm = TRUE), 2),
    samples = n()
  ) %>%
  ungroup() %>%
  mutate(trip_type = "W_HBW_v0") %>%
  left_join(rule_tbl, by = "category") %>%
  relocate(trip_type:rule, .before = category) %>%
  arrange(rate) %>%
  select(-category)
rate_list$W_HBW_v0 <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

## vi

```{r}
est_tbl <- joined %>%
  select(
    W_HBW, WeightCaliper,
    HHSize:VehSufficiency,
    -w_access, -HHSize
  ) %>%
  filter(VehSufficiency == "vi")

tree <- rpart(
  W_HBW ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 5,
  cp = .001,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, 6)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rule_tbl <- create_rule_table(tree, "vi") %>%
  select(-rate)
rate_tbl <- est_tbl %>%
  mutate(category = tree$where,) %>%
  group_by(category) %>%
  dplyr::summarize(
    rate = round(weighted.mean(W_HBW, WeightCaliper), 4),
    stdev = round(sd(W_HBW, na.rm = TRUE), 2),
    samples = n()
  ) %>%
  ungroup() %>%
  mutate(trip_type = "W_HBW_vi") %>%
  left_join(rule_tbl, by = "category") %>%
  relocate(trip_type:rule, .before = category) %>%
  arrange(rate) %>%
  select(-category)
rate_list$W_HBW_vi <- rate_tbl
```

## vs

```{r}
est_tbl <- joined %>%
  select(
    W_HBW, WeightCaliper,
    HHSize:VehSufficiency
  ) %>%
  mutate(over_60 = ifelse(Age >= 61, "Yes", "No")) %>%
  select(-Age) %>%
  filter(VehSufficiency == "vs")

tree <- rpart(
  W_HBW ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 5,
  cp = .001,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, c(29, 31))
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rule_tbl <- create_rule_table(tree, "vs") %>%
  select(-rate)
rate_tbl <- est_tbl %>%
  mutate(category = tree$where,) %>%
  group_by(category) %>%
  dplyr::summarize(
    rate = round(weighted.mean(W_HBW, WeightCaliper), 4),
    stdev = round(sd(W_HBW, na.rm = TRUE), 2),
    samples = n()
  ) %>%
  ungroup() %>%
  mutate(
    trip_type = "W_HBW_vs",
    rate = round(rate, 4)
  ) %>%
  left_join(rule_tbl, by = "category") %>%
  relocate(trip_type:rule, .before = category) %>%
  arrange(rate) %>%
  select(-category)
rate_list$W_HBW_vs <- rate_tbl
```

# W_HBO by Vehicle Sufficiency {.tabset}

## v0

```{r}
est_tbl <- joined %>%
  select(
    W_HBO, WeightCaliper,
    HHSize:VehSufficiency
  ) %>%
  filter(VehSufficiency == "v0")
  

tree <- rpart(
  W_HBO ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 2,
  cp = .001,
  minbucket = 30,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
# tree <- snip.rpart(tree, 2)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rule_tbl <- create_rule_table(tree, "v0") %>%
  select(-rate)
rate_tbl <- est_tbl %>%
  mutate(category = tree$where,) %>%
  group_by(category) %>%
  dplyr::summarize(
    rate = round(weighted.mean(W_HBO, WeightCaliper), 4),
    stdev = round(sd(W_HBO, na.rm = TRUE), 2),
    samples = n()
  ) %>%
  ungroup() %>%
  mutate(trip_type = "W_HBO_v0") %>%
  left_join(rule_tbl, by = "category") %>%
  relocate(trip_type:rule, .before = category) %>%
  arrange(rate) %>%
  select(-category)
rate_list$W_HBO_v0 <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

## vi

```{r}
est_tbl <- joined %>%
  select(
    W_HBO, WeightCaliper,
    HHSize:VehSufficiency,
    -w_access
  ) %>%
  filter(VehSufficiency == "vi")
  

tree <- rpart(
  W_HBO ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 5,
  cp = .01,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, 29)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rule_tbl <- create_rule_table(tree, "vi") %>%
  select(-rate)
rate_tbl <- est_tbl %>%
  mutate(category = tree$where,) %>%
  group_by(category) %>%
  dplyr::summarize(
    rate = round(weighted.mean(W_HBO, WeightCaliper), 4),
    stdev = round(sd(W_HBO, na.rm = TRUE), 2),
    samples = n()
  ) %>%
  ungroup() %>%
  mutate(trip_type = "W_HBO_vi") %>%
  left_join(rule_tbl, by = "category") %>%
  relocate(trip_type:rule, .before = category) %>%
  arrange(rate) %>%
  select(-category)
rate_list$W_HBO_vi <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

## vs

```{r}
est_tbl <- joined %>%
  select(
    W_HBO, WeightCaliper,
    HHSize:VehSufficiency
  ) %>%
  filter(VehSufficiency == "vs")
  

tree <- rpart(
  W_HBO ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 4,
  cp = .01,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
# tree <- snip.rpart(tree, 2)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rule_tbl <- create_rule_table(tree, "vs") %>%
  select(-rate)
rate_tbl <- est_tbl %>%
  mutate(category = tree$where,) %>%
  group_by(category) %>%
  dplyr::summarize(
    rate = round(weighted.mean(W_HBO, WeightCaliper), 4),
    stdev = round(sd(W_HBO, na.rm = TRUE), 2),
    samples = n()
  ) %>%
  ungroup() %>%
  mutate(trip_type = "W_HBO_vs") %>%
  left_join(rule_tbl, by = "category") %>%
  relocate(trip_type:rule, .before = category) %>%
  arrange(rate) %>%
  select(-category)
rate_list$W_HBO_vs <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

# N_HBSR by Vehicle Sufficiency {.tabset}

## v0

```{r}
est_tbl <- joined %>%
  select(
    N_HBSR, WeightCaliper,
    HHSize:VehSufficiency
  ) %>%
  filter(VehSufficiency == "v0")
  

tree <- rpart(
  N_HBSR ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 5,
  cp = .02,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, 6)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rule_tbl <- create_rule_table(tree, "v0") %>%
  select(-rate)
rate_tbl <- est_tbl %>%
  mutate(category = tree$where,) %>%
  group_by(category) %>%
  dplyr::summarize(
    rate = round(weighted.mean(N_HBSR, WeightCaliper), 4),
    stdev = round(sd(N_HBSR, na.rm = TRUE), 2),
    samples = n()
  ) %>%
  ungroup() %>%
  mutate(trip_type = "N_HBSR_v0") %>%
  left_join(rule_tbl, by = "category") %>%
  relocate(trip_type:rule, .before = category) %>%
  arrange(rate) %>%
  select(-category)
rate_list$N_HBSR_v0 <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

## vi

```{r}
est_tbl <- joined %>%
  select(
    N_HBSR, WeightCaliper,
    HHSize:VehSufficiency,
    -t_access
  ) %>%
  filter(VehSufficiency == "vi")
  

tree <- rpart(
  N_HBSR ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 4,
  cp = .01,
  minbucket = 59,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, c(6, 14))
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rule_tbl <- create_rule_table(tree, "vi") %>%
  select(-rate)
rate_tbl <- est_tbl %>%
  mutate(category = tree$where,) %>%
  group_by(category) %>%
  dplyr::summarize(
    rate = round(weighted.mean(N_HBSR, WeightCaliper), 4),
    stdev = round(sd(N_HBSR, na.rm = TRUE), 2),
    samples = n()
  ) %>%
  ungroup() %>%
  mutate(trip_type = "N_HBSR_vi") %>%
  left_join(rule_tbl, by = "category") %>%
  relocate(trip_type:rule, .before = category) %>%
  arrange(rate) %>%
  select(-category)
rate_list$N_HBSR_vi <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

## vs

```{r}
est_tbl <- joined %>%
  mutate(IncPerCapLt80 = ifelse(IncPerCapita < 80000, "Yes", "No")) %>%
  select(
    N_HBSR, WeightCaliper,
    HHSize:VehSufficiency,
    -HHSize, -Age, -HHIncCat, -IncPerCapita,
    IncPerCapLt80,
    -Vehicles, -t_access, -g_access
  ) %>%
  filter(VehSufficiency == "vs")
  

tree <- rpart(
  N_HBSR ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 4,
  cp = .007,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, 13)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rule_tbl <- create_rule_table(tree, "vs") %>%
  select(-rate)
rate_tbl <- est_tbl %>%
  mutate(category = tree$where,) %>%
  group_by(category) %>%
  dplyr::summarize(
    rate = round(weighted.mean(N_HBSR, WeightCaliper), 4),
    stdev = round(sd(N_HBSR, na.rm = TRUE), 2),
    samples = n()
  ) %>%
  ungroup() %>%
  mutate(trip_type = "N_HBSR_vs") %>%
  left_join(rule_tbl, by = "category") %>%
  relocate(trip_type:rule, .before = category) %>%
  arrange(rate) %>%
  select(-category)
rate_list$N_HBSR_vs <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

# N_HBO by Vehicle Sufficiency {.tabset}

## v0

```{r}
est_tbl <- joined %>%
  select(
    N_HBO, WeightCaliper,
    HHSize:VehSufficiency
    # -w_access
  ) %>%
  mutate(t_access_lt_61 = ifelse(t_access < 6.1, "Yes", "No")) %>%
  select(-t_access, -w_access) %>%
  filter(VehSufficiency == "v0")
  

tree <- rpart(
  N_HBO ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 4,
  cp = .01,
  minbucket = 30,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
# tree <- snip.rpart(tree, 2)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rule_tbl <- create_rule_table(tree, "v0") %>%
  select(-rate)
rate_tbl <- est_tbl %>%
  mutate(category = tree$where,) %>%
  group_by(category) %>%
  dplyr::summarize(
    rate = round(weighted.mean(N_HBO, WeightCaliper), 4),
    stdev = round(sd(N_HBO, na.rm = TRUE), 2),
    samples = n()
  ) %>%
  ungroup() %>%
  mutate(trip_type = "N_HBO_v0") %>%
  left_join(rule_tbl, by = "category") %>%
  relocate(trip_type:rule, .before = category) %>%
  arrange(rate) %>%
  select(-category)
rate_list$N_HBO_v0 <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

## vi

```{r}
est_tbl <- joined %>%
  select(
    N_HBO, WeightCaliper,
    HHSize:VehSufficiency,
    -IncPerCapita, -HHIncCat
  ) %>%
  filter(VehSufficiency == "vi")
  

tree <- rpart(
  N_HBO ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 4,
  cp = .01,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
# tree <- snip.rpart(tree, 2)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rule_tbl <- create_rule_table(tree, "vi") %>%
  select(-rate)
rate_tbl <- est_tbl %>%
  mutate(category = tree$where,) %>%
  group_by(category) %>%
  dplyr::summarize(
    rate = round(weighted.mean(N_HBO, WeightCaliper), 4),
    stdev = round(sd(N_HBO, na.rm = TRUE), 2),
    samples = n()
  ) %>%
  ungroup() %>%
  mutate(trip_type = "N_HBO_vi") %>%
  left_join(rule_tbl, by = "category") %>%
  relocate(trip_type:rule, .before = category) %>%
  arrange(rate) %>%
  select(-category)
rate_list$N_HBO_vi <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

## vs

```{r}
est_tbl <- joined %>%
  select(
    N_HBO, WeightCaliper,
    HHSize:VehSufficiency
  ) %>%
  filter(VehSufficiency == "vs")
  

tree <- rpart(
  N_HBO ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 4,
  cp = .01,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
# tree <- snip.rpart(tree, 2)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rule_tbl <- create_rule_table(tree, "vs") %>%
  select(-rate)
rate_tbl <- est_tbl %>%
  mutate(category = tree$where,) %>%
  group_by(category) %>%
  dplyr::summarize(
    rate = round(weighted.mean(N_HBO, WeightCaliper), 4),
    stdev = round(sd(N_HBO, na.rm = TRUE), 2),
    samples = n()
  ) %>%
  ungroup() %>%
  mutate(trip_type = "N_HBO_vs") %>%
  left_join(rule_tbl, by = "category") %>%
  relocate(trip_type:rule, .before = category) %>%
  arrange(rate) %>%
  select(-category)
rate_list$N_HBO_vs <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

# N_HBSCH

```{r}
est_tbl <- joined %>%
  select(
    N_HBSCH, WeightCaliper,
    HHSize:VehSufficiency, -HHSize, -HHKids, -t_access
  )

tree <- rpart(
  N_HBSCH ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 4,
  cp = .001,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, 15)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rule_tbl <- create_rule_table(tree) %>%
  select(-rate)
rate_tbl <- est_tbl %>%
  mutate(category = tree$where,) %>%
  group_by(category) %>%
  dplyr::summarize(
    rate = round(weighted.mean(N_HBSCH, WeightCaliper), 4),
    stdev = round(sd(N_HBSCH, na.rm = TRUE), 2),
    samples = n()
  ) %>%
  ungroup() %>%
  mutate(trip_type = "N_HBSCH") %>%
  left_join(rule_tbl, by = "category") %>%
  relocate(trip_type:rule, .before = category) %>%
  arrange(rate) %>%
  select(-category)
rate_list$N_HBSCH <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

# N_HBSHP by Vehicle Sufficiency {.tabset}

## v0

```{r}
est_tbl <- joined %>%
  select(
    N_HBSHP, WeightCaliper,
    HHSize:VehSufficiency, -IncPerCapita, -w_access, -t_access
  ) %>%
  filter(VehSufficiency == "v0")
  

tree <- rpart(
  N_HBSHP ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 5,
  cp = .001,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, 12)
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rule_tbl <- create_rule_table(tree, "v0") %>%
  select(-rate)
rate_tbl <- est_tbl %>%
  mutate(category = tree$where,) %>%
  group_by(category) %>%
  dplyr::summarize(
    rate = round(weighted.mean(N_HBSHP, WeightCaliper), 4),
    stdev = round(sd(N_HBSHP, na.rm = TRUE), 2),
    samples = n()
  ) %>%
  ungroup() %>%
  mutate(trip_type = "N_HBSHP_v0") %>%
  left_join(rule_tbl, by = "category") %>%
  relocate(trip_type:rule, .before = category) %>%
  arrange(rate) %>%
  select(-category)
rate_list$N_HBSHP_v0 <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

## vi

```{r}
est_tbl <- joined %>%
  select(
    N_HBSHP, WeightCaliper,
    HHSize:VehSufficiency, -IncPerCapita,
    -HHAdults, -contains("access")
  ) %>%
  filter(VehSufficiency == "vi")
  

tree <- rpart(
  N_HBSHP ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 4,
  cp = .001,
  minbucket = 30,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, c(9, 10, 11))
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rule_tbl <- create_rule_table(tree, "vi") %>%
  select(-rate)
rate_tbl <- est_tbl %>%
  mutate(category = tree$where,) %>%
  group_by(category) %>%
  dplyr::summarize(
    rate = round(weighted.mean(N_HBSHP, WeightCaliper), 4),
    stdev = round(sd(N_HBSHP, na.rm = TRUE), 2),
    samples = n()
  ) %>%
  ungroup() %>%
  mutate(trip_type = "N_HBSHP_vi") %>%
  left_join(rule_tbl, by = "category") %>%
  relocate(trip_type:rule, .before = category) %>%
  arrange(rate) %>%
  select(-category)
rate_list$N_HBSHP_vi <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```

## vs

```{r}
est_tbl <- joined %>%
  select(
    N_HBSHP, WeightCaliper,
    HHSize:VehSufficiency,
    -Vehicles, -w_access, -IncPerCapita, -t_access
  ) %>%
  filter(VehSufficiency == "vs")
  

tree <- rpart(
  N_HBSHP ~ . - WeightCaliper,
  data = est_tbl, 
  method = "anova",
  maxdepth = 5,
  cp = .007,
  minbucket = 40,
  weights = est_tbl$WeightCaliper
)

# Snip to remove specific nodes from the tree
# summary(tree) # to find node numbers
tree <- snip.rpart(tree, c(9, 5))
rpart.plot(tree, type = 1, under = TRUE, extra = 1)

rule_tbl <- create_rule_table(tree, "vs") %>%
  select(-rate)
rate_tbl <- est_tbl %>%
  mutate(category = tree$where,) %>%
  group_by(category) %>%
  dplyr::summarize(
    rate = round(weighted.mean(N_HBSHP, WeightCaliper), 4),
    stdev = round(sd(N_HBSHP, na.rm = TRUE), 2),
    samples = n()
  ) %>%
  ungroup() %>%
  mutate(trip_type = "N_HBSHP_vs") %>%
  left_join(rule_tbl, by = "category") %>%
  relocate(trip_type:rule, .before = category) %>%
  arrange(rate) %>%
  select(-category)
rate_list$N_HBSHP_vs <- rate_tbl

# check r^2 at household level
# check <- joined %>%
#   mutate(
#     y = W_HBW,
#     y_hat = predict(tree, est_tbl)
#   ) %>%
  # group_by(Sample_Number) %>%
  # summarize(
  #   y = sum(y),
  #   y_hat = sum(y_hat)
  # )
# r_sq <- round(cor(check$y, check$y_hat)^2, 3)
# rsq.rpart(tree)
```


```{r, eval=FALSE}
out_df <- bind_rows(rate_list) %>%
  mutate(
    rule = gsub("VehSufficiency = v0,vi", "VehSufficiency <> vs", rule, fixed = TRUE),
    rule = gsub("VehSufficiency = v0,vs", "VehSufficiency <> vi", rule, fixed = TRUE),
    rule = gsub("VehSufficiency", "auto_suff", rule, fixed = TRUE),
    rule = gsub(" vi", " 'vi'", rule, fixed = TRUE),
    rule = gsub(" vs", " 'vs'", rule, fixed = TRUE),
    rule = gsub(" v0", " 'v0'", rule, fixed = TRUE),
    rule = gsub("= No", "= 0", rule, fixed = TRUE),
    rule = gsub("= Yes", "= 1", rule, fixed = TRUE),
    rule = gsub("= Male", "= 1", rule, fixed = TRUE),
    rule = gsub("= Female", "= 2", rule, fixed = TRUE),
  )
write_csv(out_df, "data/output/production_trees/production_rates.csv")

# write_csv(
#   calib_factors,
#   "../master/resident/generation/calibration_factors.csv"
# )
```